---
title: "**PREY EXHIBIT LESS RISKY BUT MORE VARIABLE BEHAVIOUR IN HIGH PREDATION ENVIRONMENTS**"
subtitle: "**Supplementary Material**"
written by: "**EIF Wooster, EJ Lundgren, K Wrensford, KM Gaynor & KL Laskowski**"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    theme: simplex
    number_sections: true
    code_folding: hide
    self_contained: false
  
---

------------------------------------------------------------------------
## Setting up

### Loading Packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,    
                      message = FALSE)
#knitr::opts_chunk$set(eval = FALSE)
```

```{r, echo = T}

# Loading Packages

pacman::p_load(devtools, 
               tidyverse, 
               metafor, 
               patchwork, 
               R.rsp, orchaRd,
               emmeans,
               metafor,
               orchaRd,
               stringr,
               dplyr,
               broom,
               tidyverse,
               here,
               broom,
               ggplot2,
               viridis,
               data.table,
               Matrix, 
               matrixcalc,
               ape,
               multcomp,
               taxize,
               cowplot,
               kableExtra,
               Pandoc,
               magrittr
) 



```


### Reading in Dataset and Phylogentic Tree

```{r, echo = T, results = 'hide'}

# Reading in Data
dat <- read.csv(here("Data/Analysis_ready.csv")) 

#change a few specie names for analysis

dat$Species <- gsub("Equus_quagga","Equus_burchellii_quagga", dat$Species)
dat$Species <- gsub("Macropus_eugenii","Notamacropus_eugenii",  dat$Species)
dat$Species <- gsub("Hyla_versicolor", "Dryophytes_versicolor",  dat$Species)
dat$Species <- gsub("Bufo_americanus", "Anaxyrus_americanus",  dat$Species)
dat$Species <- gsub( "Gambusiaa_hubbsi","Gambusia_hubbsi",  dat$Species)
dat$Species <- gsub( "Gambusia_hubbsi ","Gambusia_hubbsi",  dat$Species)
dat$Species <- gsub( "Brachyraphis_episcopi","Brachyrhaphis_episcopi",  dat$Species)

# Reading in tree

tree1 <- read.tree(here("Tree/Tree_risk.tre")) 

# Getting branch length and correlation matrix

tree1b <- compute.brlen(tree1)
cor1 <-  vcv(tree1b, corr=T)

# checking the match 

setdiff(dat$Species, tree1$tip.label)
setdiff(tree1$tip.label, dat$Species)

# creating non-phylo columns - has to be tree column 

dat$Species2 <- dat$Species

```

### Dataset  
```{r, echo = T}

# making a scrollable table
kable(dat, "html", escape = FALSE) %>% kable_styling("striped", position = "left", full_width = TRUE) %>% 
  column_spec(ncol(dat), width_min = "5in") %>% 
  scroll_box(width = "100%", 
    height = "500px")

```

### Converting error and adding columns for RE
```{r, echo = T, results = 'hide'}
### functions for converting error ------

# Function for converting CI's to SD

me_to_sd <- function(me, n, ci_level = 0.95) {
  alpha <- 1 - ci_level
  tcrit <- qt(1 - alpha/2, df = n - 1)
  me * sqrt(n) / tcrit
}

# Function for converting SE to SD

se_to_sd <- function(se, n) {
  # se: standard error
  # n: sample size
  # return: standard deviation
  sd_value <- se * sqrt(n)
  return(sd_value)
}

### Converting error types

unique(dat$Error_type)

### >>> Standard deviation --------

SD <- filter(dat, Error_type == "SD")

SD$SD_cont_trans <- SD$SD_control

SD$SD_exp_trans <- SD$SD_exp

### >>> 95% CI's --------

CI <- filter(dat, Error_type == "CI")

CI$SD_cont_trans <- me_to_sd(CI$SD_control, CI$n_control)

CI$SD_exp_trans <- me_to_sd(CI$SD_exp, CI$n_exp)

### >>> Standard error --------

SE <- filter(dat, Error_type == "SE")

SE$SD_cont_trans <- se_to_sd(SE$SD_control, SE$n_control)

SE$SD_exp_trans <- se_to_sd(SE$SD_exp, SE$n_exp)

# Put them back together!!

data <- rbind(SD, CI, SE)

# obs level random effect

data$Obs_ID <- factor(1:nrow(data))

# Study level random effect 

data$Firstauthor_Year <- paste(data$First_author, data$Year, sep = "_")

data$Study_ID <- factor(as.numeric(as.factor(data$Firstauthor_Year)))

```
### Calculating SMD and lnCVR effect size

```{r, echo = T}

# Calculate Hedges' g

es <- escalc(
  measure = "SMDH",     
  m1i = Mean_exp,
  sd1i = SD_exp_trans,
  n1i = n_exp,
  m2i = Mean_control,
  sd2i = SD_cont_trans,
  n2i = n_control,
  data = data
)

##now we need to invert the effect sizes the sign on "yes" on sign-inverted to ensure risk = bigger

es_inv <- filter(es, Sign_inverted == "Yes")

es_not <- filter(es, Sign_inverted == "No")

#Ok lets flip the sign back for inv

es_inv$yi <- -es_inv$yi


es2 <- rbind(es_inv, es_not)


# Calculate lnCVR

#removing negatives from the dataset
cv <- filter(data, Mean_control > 0)

cv <- filter(cv, Mean_exp > 0)

#remove methods where a value could be negative 

cv <- filter(cv, Can_be_negitive != "Yes")


# Calculate lncvr

cv <- escalc(
  measure = "CVR",     
  m1i = Mean_exp,
  sd1i = SD_exp_trans,
  n1i = n_exp,
  m2i = Mean_control,
  sd2i = SD_cont_trans,
  n2i = n_control,
  data = cv
)

cv$Obs_ID <- factor(1:nrow(cv))

```

## Meta-analysis SMD  {.tabset}

```{r, echo = T}

mod.overall.simple1 <- rma.mv(yi = yi, V = vi, 
                              random = list(~1 | Study_ID / Obs_ID, 
                                            ~1 | Species2 
                              ), 
                              data =  es2,
                              # control = list(optimizer="BFGS"),
                              test = "t",
                              sparse = TRUE)

summary(mod.overall.simple1)

overall <- orchard_plot(mod.overall.simple1, xlab = "Difference in risk-taking (SMDH)", group = "Study_ID",
                        angle = 0, col = "#eea196")

overall

round(i2_ml(mod.overall.simple1), 2)

```

## Meta-regression SMD:Uni-moderator {.tabset}

### Behaviour type

```{r, echo = T}

es2$Behaviour_type <- factor(
  es2$Behaviour_type,
  levels = sort(unique(es2$Behaviour_type), decreasing = TRUE)
)


mod.behav <- rma.mv(yi = yi, V = vi,
                      random = list(~1 | Study_ID / Obs_ID, 
                                    ~1 | Species2
                      ), 
                      data =  es2,
                      mods = ~ Behaviour_type -1,
                      test = "t",
                      sparse = TRUE)
        


summary(mod.behav)


behav <- orchard_plot(mod.behav, mod = "Behaviour_type", 
                      xlab = "Difference in risk-taking behaviour (SMDH)", 
                      group = "Study_ID",
                      angle = 0) +
  scale_fill_manual(values = rep("#eea196", 7)) +
  scale_color_manual(values = rep("#eea196", 7))

behav

```

### Sex

```{r, echo = T}
ses <- filter(es2, Sex != "N/A")

mod.sex <- rma.mv(yi = yi, V = vi,
                  random = list(~1 | Study_ID / Obs_ID, 
                                ~1 | Species2
                  ), 
                    data =  ses,
                    mods = ~ Sex -1,
                    test = "t",
                    sparse = TRUE)

summary(mod.sex)

sexplot <- orchard_plot(mod.sex, xlab = "Difference in risk-taking behaviour (SMDH)", 
                           group = "Study_ID", mod = "Sex")

sexplot

```

### Real vs simulated predation

```{r, echo = T}

es2$Real_predator <- as.factor(es2$Real_predator)


mod.rp <- rma.mv(yi = yi, V = vi,
                 random = list(~1 | Study_ID / Obs_ID, 
                               ~1 | Species2
                 ), 
                 data =  es2, 
                 mods = ~ Real_predator -1,
                 test = "t",
                 sparse = TRUE)

summary(mod.rp)

rp <- orchard_plot(mod.rp, mod = "Real_predator", 
                   xlab = "Difference in risk-taking behaviour (SMDH)", 
                   group = "Study_ID",
                   angle = 0) +
  scale_fill_manual(values = rep("#eea196", 7)) +
  scale_color_manual(values = rep("#eea196", 7))

rp

```

### Low vs no predator as control

```{r, echo = T}

mod.ln <- rma.mv(yi = yi, V = vi,
                 random = list(~1 | Study_ID / Obs_ID, 
                               ~1 | Species2
                 ), 
                 data =  es2, 
                 mods = ~ Low_or_no_pred -1,
                 test = "t",
                 sparse = TRUE)

summary(mod.ln)

ln <- orchard_plot(mod.ln, mod = "Low_or_no_pred", 
                   xlab = "Difference in risk-taking behaviour (SMDH)", 
                   group = "Study_ID",
                   angle = 0) +
  scale_fill_manual(values = rep("#eea196", 7)) +
  scale_color_manual(values = rep("#eea196", 7))

ln

```

### adult vs juvenille

```{r, echo = T}
es2$Adult <- factor(es2$Adult, 
                              levels = c("Adult", "Juvenille", "Both"))

mod.aj <- rma.mv(yi = yi, V = vi,
                 random = list(~1 | Study_ID / Obs_ID, 
                               ~1 | Species2 
                 ), 
                 data =  es2, 
                 mods = ~ Adult -1,
                 test = "t",
                 sparse = TRUE)

summary(mod.aj)

aj <- orchard_plot(mod.aj, mod = "Adult", 
                   xlab = "Difference in risk-taking behaviour (SMDH)", 
                   group = "Study_ID",
                   angle = 0) +
  scale_fill_manual(values = rep("#eea196", 7)) +
  scale_color_manual(values = rep("#eea196", 7))

aj

```

### Comparision type 

```{r, echo = T}
mod.com <- rma.mv(yi = yi, V = vi,
                  random = list(~1 | Study_ID / Obs_ID, 
                                ~1 | Species2
                  ), 
                  data =  es2, 
                 mods = ~ Comparison_type -1,
                 test = "t",
                 sparse = TRUE)

summary(mod.com)

ct <- orchard_plot(mod.com, mod = "Comparison_type", 
                   xlab = "Difference in risk-taking behaviour (SMDH)", 
                   group = "Study_ID",
                   angle = 0) +
  scale_fill_manual(values = rep("#eea196", 7)) +
  scale_color_manual(values = rep("#eea196", 7))

ct

```

### Class

```{r, echo = T}
s2 <- es2 %>%
  mutate(Class = recode(Class,
                        "Actinopterygii" = "Ray-finned fishes",
                        "Amphibia" = "Amphibians",
                        "Mammalia" = "Mammals",
                        "Reptilia" = "Reptiles",
                        "Malacostraca" = "Crustaceans",
                        "Chondrichthyes" = "Cartilaginous fishes",
                        "Aves" = "Birds"
  ))

es2$Class <- factor(
  es2$Class,
  levels = sort(unique(es2$Class), decreasing = TRUE)
)


mod.class <- rma.mv(yi = yi, V = vi,
                    random = list(~1 | Study_ID / Obs_ID, 
                                  ~1 | Species2
                    ), 
                  data =  es2,
                  mods = ~ Class -1,
                  test = "t",
                  sparse = TRUE)
                

summary(mod.class)

class <- orchard_plot(mod.class, mod = "Class", xlab = "Difference in risk-taking (SMDH)", group = "Study_ID",
                      angle = 0) + 
  scale_fill_manual(values = rep("#eea196", 7)) +
  scale_color_manual(values = rep("#eea196", 7))
class   

```

## Meta-analysis lnCVR  {.tabset}

```{r, echo = T}
mod.overall_cv <- rma.mv(yi = yi, V = vi,
                         random = list(~1 | Study_ID / Obs_ID), 
                         data =  cv,
                         test = "t",
                         sparse = TRUE)


summary(mod.overall_cv)


overall_cv <- orchard_plot(mod.overall_cv, xlab = "Difference in heterogeneity (lnCVR)", group = "Study_ID",
                           angle = 0)
overall_cv <- overall_cv + scale_fill_manual(values = "#989aae") + scale_color_manual(values = "#989aae") 

overall_cv

```

## Meta-regression lnCVR:Uni-moderator {.tabset}

### Behaviour type lnCVR

```{r, echo = T}
cv$Behaviour_type <- factor(
  cv$Behaviour_type,
  levels = sort(unique(cv$Behaviour_type), decreasing = TRUE)
)

mod.behav_cv <- rma.mv(yi = yi, V = vi,
                       random = list(~1 | Study_ID / Obs_ID), 
                       data =  cv,
                       mods = ~ Behaviour_type -1,
                       test = "t",
                       sparse = TRUE)

summary(mod.behav_cv)

behav.cv <- orchard_plot(mod.behav_cv, mod = "Behaviour_type", 
                      xlab = "Difference in heterogeneity (lnCVR)", 
                      group = "Study_ID",
                      angle = 0) 

behav.cv <- behav.cv +
  scale_fill_manual(values = rep("#989aae", 7)) +
  scale_color_manual(values = rep("#989aae", 7))
behav.cv

```


### Sex lnCVR

```{r, echo = T}
ses_cv <- filter(cv, Sex != "N/A")


mod.sex.cv <- rma.mv(yi = yi, V = vi,
                   random = list(~1 | Study_ID / Obs_ID), 
                     data =  ses_cv,
                     mods = ~ Sex -1,
                     test = "t",
                     sparse = TRUE)

summary(mod.sex.cv) 

behav.cv <- orchard_plot(mod.sex.cv, mod = "Sex", 
                      xlab = "Difference in heterogeneity (lnCVR)", 
                      group = "Study_ID",
                      angle = 0) 

behav.cv <- behav.cv +
  scale_fill_manual(values = rep("#989aae", 7)) +
  scale_color_manual(values = rep("#989aae", 7))
behav.cv

```

### Real vs simulated predation lnCVR

```{r, echo = T}
cv$Real_predator <- as.factor(cv$Real_predator)
mod.rp.cv <- rma.mv(yi = yi, V = vi,
                    random = list(~1 | Study_ID / Obs_ID), 
                    data =  cv,
                    mods = ~ Real_predator -1,
                    test = "t",
                    sparse = TRUE)

summary(mod.rp.cv)

rp.cv <- orchard_plot(mod.rp.cv, mod = "Real_predator", 
                   xlab = "Difference in heterogeneity (lnCVR)", 
                   group = "Study_ID",
                   angle = 0) +
  scale_fill_manual(values = rep("#989aae", 7)) +
  scale_color_manual(values = rep("#989aae", 7))

rp.cv

```

### Low vs no predator as control lnCVR

```{r, echo = T}
mod.ln.cv <- rma.mv(yi = yi, V = vi,
                    random = list(~1 | Study_ID / Obs_ID), 
                    data =  cv,
                    mods = ~ Low_or_no_pred -1,
                    test = "t",
                    sparse = TRUE)

summary(mod.ln.cv)


ln.cv <- orchard_plot(mod.ln.cv, mod = "Low_or_no_pred", 
                      xlab = "Difference in heterogeneity (lnCVR)", 
                      group = "Study_ID",
                      angle = 0) +
  scale_fill_manual(values = rep("#989aae", 7)) +
  scale_color_manual(values = rep("#989aae", 7))

ln.cv

```

### Adult vs juvenille lnCVR

```{r, echo = T}
mod.ln.cv <- rma.mv(yi = yi, V = vi,
                    random = list(~1 | Study_ID / Obs_ID), 
                    data =  cv,
                    mods = ~ Low_or_no_pred -1,
                    test = "t",
                    sparse = TRUE)

summary(mod.ln.cv)


ln.cv <- orchard_plot(mod.ln.cv, mod = "Low_or_no_pred", 
                      xlab = "Difference in heterogeneity (lnCVR)", 
                      group = "Study_ID",
                      angle = 0) +
  scale_fill_manual(values = rep("#989aae", 7)) +
  scale_color_manual(values = rep("#989aae", 7))

ln.cv

```


### Comparison type lnCVR

```{r, echo = T}
mod.com.cv <- rma.mv(yi = yi, V = vi,
                     random = list(~1 | Study_ID / Obs_ID), 
                     data =  cv,
                     mods = ~ Comparison_type -1,
                     test = "t",
                     sparse = TRUE)

summary(mod.com.cv)


ct.cv <- orchard_plot(mod.com.cv, mod = "Comparison_type", 
                      xlab = "Difference in heterogeneity (lnCVR)", 
                      group = "Study_ID",
                      angle = 0) +
  scale_fill_manual(values = rep("#989aae", 7)) +
  scale_color_manual(values = rep("#989aae", 7))

ct.cv 
```

### Class lnCVR

```{r, echo = T}
cv <- cv %>%
  mutate(Class = recode(Class,
                        "Actinopterygii" = "Ray-finned fishes",
                        "Amphibia" = "Amphibians",
                        "Mammalia" = "Mammals",
                        "Reptilia" = "Reptiles",
                        "Malacostraca" = "Crustaceans",
                        "Chondrichthyes" = "Cartilaginous fishes",
                        "Aves" = "Birds"
  ))



cv$Class <- factor(
  cv$Class,
  levels = sort(unique(cv$Class), decreasing = TRUE)
)

mod.class.cv <- rma.mv(yi = yi, V = vi,
                       random = list(~1 | Study_ID / Obs_ID), 
                       data =  cv,
                       mods = ~ Class -1,
                       test = "t",
                       sparse = TRUE)

summary(mod.class.cv)

class.cv <- orchard_plot(mod.class.cv, mod = "Class", xlab = "Heterogeneity in risk-taking (lnCVR)", group = "Study_ID",
                      angle = 0) + 
  scale_fill_manual(values = rep("#989aae", 7)) +
  scale_color_manual(values = rep("#989aae", 7))
class.cv  
```



## Publication bias {.tabset}

### Decline effect for difference

```{r, echo = T}

mod.mad <- rma.mv(yi = yi, V = vi,
                  random = list(~1 | Study_ID / Obs_ID, 
                                ~1 | Species2
                  ), 
                  data =  es2,
                          mods = ~ Year,
                          test = "t",
                          sparse = TRUE)


summary(mod.mad) 

```

### Eggers regression + funnel

```{r, echo = T}

es2$effectN <- (4* es2$n_control * es2$n_exp) / (es2$n_control + es2$n_exp)
es2$sqeffectN <- sqrt(1/es2$effectN)

#' [EJL Changed:]
mod.egg <- rma.mv(yi = yi, V = vi,
                  random = list(~1 | Study_ID / Obs_ID, #' [EJL changed]
                                #~1 | Species, # phylo effect 
                                ~1 | Species2#, # non-phylo effect 
                  ), #' [EJL Change]
                  # ~1 | Species, # phylo effect 
                  # ~1 | Species2),  
                  data =  es2,
                  mods = ~ sqeffectN,
                  test = "t",
                  sparse = TRUE)

summary(mod.egg)


funnel(mod.egg, 
       yaxis="seinv",
       xlab = "Standardized residuals",
       ylab = "Precision (inverse of SE)",
       
)
```

### Leave one out

```{r, echo = T}

dat <- es2 %>%
  mutate(leave_out = paste(First_author, Year, sep = "_"))
dat$leave_out <- as.factor(dat$leave_out)

  LeaveOneOut_effectsize <- list()
  for (i in 1:length(levels(dat$leave_out))) {
    temp_dat <- dat %>%
      dplyr::filter(leave_out != levels(dat$leave_out)[i])
    
    LeaveOneOut_effectsize[[i]] <- rma.mv(yi = yi,
                                          V = vi,
                                          random = list(~1 | Study_ID / Obs_ID,
                                                        ~1 | Species2),
                                 
                                          test = "t",
                                          method = "REML",
                                          sparse = TRUE,
                                          data = temp_dat)
  }
  
  # function for extracting est, ci.lb, and ci.ub from all models
  est.func <- function(model) {
    df <- data.frame(est = model$b, lower = model$ci.lb, upper = model$ci.ub)
    return(df)
  } 
    
  # form data frame
  MA_oneout <- lapply(LeaveOneOut_effectsize, function(x) est.func(x)) %>%
    bind_rows() %>%
    mutate(left_out = levels(dat$leave_out))
  
  # preserve factor order for ggplot
  MA_oneout$left_out <- factor(MA_oneout$left_out, levels = MA_oneout$left_out)
  
  # save the runs
  saveRDS(MA_oneout, here("R", "MA_oneout.RDS"))
  


# plotting
leaveoneout <- ggplot(MA_oneout) +
  geom_hline(yintercept = 0, lty = 2, lwd = 1) +
  geom_hline(yintercept = mod.overall.simple1$ci.lb, lty = 3, lwd = 0.75, colour = "black") +
  geom_hline(yintercept = mod.overall.simple1$b,     lty = 1, lwd = 0.75, colour = "black") +
  geom_hline(yintercept = mod.overall.simple1$ci.ub, lty = 3, lwd = 0.75, colour = "black") +
  geom_pointrange(aes(x = left_out, y = est, ymin = lower, ymax = upper)) +
  xlab("Study left out") +
  ylab("Hedge' g") +
  coord_flip() +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(size = 6)
  )

leaveoneout



```

### Sensivity analysis for lnCVR as representative sample

```{r, echo = T}
sens.es <- filter(es2, Mean_control > 0)

sens.es <- filter(sens.es, Mean_exp > 0)

sens.es <- filter(sens.es, Can_be_negitive != "Yes")

sens.es$Obs_ID <- factor(1:nrow(sens.es))

mod.behav.sens <- rma.mv(yi = yi, V = vi,
                         random = list(~1 | Study_ID / Obs_ID, 
                                       ~1 | Species2
                         ), 
                    data =  sens.es,
                    test = "t",
                    sparse = TRUE)

summary(mod.behav.sens)
```


### Sensivity analysis for lnCVR as representative sample
```{r, echo = T}
sens.es.pca <- filter(es2, Metric != "PCA values")

mod.behav.sens.pca <- rma.mv(yi = yi, V = vi,
                         random = list(~1 | Study_ID / Obs_ID, 
                                       ~1 | Species2
                         ), 
                         data =  sens.es.pca,
                         test = "t",
                         sparse = TRUE)

summary(mod.behav.sens.pca)
```

## R Session Information

```{r, echo = T}
library(sessioninfo)
session_info()
